A step towards continuous Contests

### 2016-02-16 ms.tex
cts contest theory merged into ms, plots of cts in ms

Within the immature phase, we assume males grow with rate
\begin{equation} \label{eq:growth}
\frac{{\rm d} m}{{\rm d} t} = a \, m^ b
\end{equation}
where $a$ is mass-based rate of growth. Integrating eq. \ref{eq:growth}, the size of individuals at time $t$, having started growing at time $0$ is given by
\begin{equation} \label{eq:growth}
m(t) = \left(m(0)^{1-b} + t a(1-b)\right)^{\frac1{1-b}},
\end{equation}

@@

Finally, we assume males transition from immature $\rightarrow$ mature according to logistic probability function
\begin{equation}\label{eq:allocation}
Pr({\rm mature} \, | m) = \frac1{1 + \exp\left(\alpha \left(1 - m / m_{\rm mat}\right)\right)},
\end{equation}
where $m_{\rm mat}$ is the size at which 50\% of the population is mature and $\alpha$ determines the width of the curve.

@@

\clearpage
\subsection{Contest Algorithm}
See Fig. \ref{fig:contests}.
For a defender $x_{def}$ vs an attacker $x_{atk}$, there are 3 (currently 2) stages:
The first stage is the display stage, the second phase is the fight stage.
each individual $i$ has mass $m$. At the time of the contest it also has an amount of energy $E_{current}$.  
The aggression of the individual is a function of $E_{current}$:

\begin{equation}\label{eq:aggression}
a_i = f(E_{current})=\frac{L}{1+e^{(-K(E_{current}-E_{o})}}
\end{equation}

$L$ is a scaling parameter.

Contests can be escalated to different degrees with increasing energy costs for both participants. 
Either participant may abandon the contest at any moment, they do this by choosing not to escalate.
We assume that there are energetic costs, but that there is no time cost to contests ie they occur instantaneously.

Traits $K$ and $E_{o}$ are heritable and potentially under selection.

The probability of the attacker escalating to  is:

\begin{equation}\label{eq:attacker_escalating}
Pr(\rm{escalation}) =  logit^{-1}(a_{atk} \cdot m_{atk} - m_{def})
\end{equation}

Similarly the probability of the defender escalating is:

\begin{equation}\label{eq:defender_escalating}
          Pr(\rm{escalation}) =  logit^{-1}(a_{def} \cdot m_{def} - m_{atk})
\end{equation}

The energetic cost of escalating is:
\begin{equation}\label{eq:defender_escalating}
E_{spent}=equation you guys figured out with all the \lambda 's
\end{equation}

The cost of the fighting stage is much greater than the cost of the display stage. Add a figure here to show this.

If both individuals escalate to the end of the fight, the winner is chosen.  $x_{def}$ wins with the probability $P_{def} = logit^{-1}(m_{def} - m_{atk})$
After the contest energy costs are deducted from both the winner and the loser.
If a death occurs ($E \leq 0$) then x is removed from the individual pool.

We are looking at two \'cost functions':
\begin{enumerate}
    \item A flat percentage of the losers maximum energy or...
    \item $$ E_{costs} = \frac {E_{loser max}} {1 + exp(-(\frac{m_{winner} - m_{loser}}{k} - m_0)}$$ Where the fight cost is a function of the losers mass and the winners mass, and a couple of other parameters.
\end{enumerate}

@@

    \item aggression 0 ($\beta$) called $e_0$
    \item aggression 1 ($\alpha$)

@@

\section{Results} % (fold)
\label{sec:results}

\begin{figure}[h!]
    \centering
    \includegraphics[width=8cm,height=10cm,keepaspectratio]{figures/energy_pc_plot.pdf}
    \caption{Flat costs. Percent energy expenditure of cohort at generation 499}
    \label{fig:energy pc}

    \centering
    \includegraphics[width=8cm,height=10cm,keepaspectratio]{figures/energy_total_plot.pdf}
    \caption{ Flat costs Total energy expenditure of cohort at generation 499}
    \label{fig:energy total}
\end{figure}
As patch area increases the number of fight importunities decreases,
This can probably explain the strong decrease in energy spent fighting.but I have no idea, early days yet.
\clearpage

\begin{figure}[h!]
    \centering
    \includegraphics[width=8cm,height=10cm,keepaspectratio]{figures/energy_pc_plot_var.pdf}
    \caption{Variable costs. Percent energy expenditure of cohort at generation 499}
    \label{fig:var energy pc}
    \centering
    \includegraphics[width=8cm,height=10cm,keepaspectratio]{figures/energy_total_plot_var.pdf}
    \caption{Variable costs, Total energy expenditure of cohort at generation 499}
    \label{fig:var energy total}
\end{figure}

The results for variable costs are different, why? from here on all graphs are from the flat cost data...

\begin{figure}[h!]
    \centering
    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/k_density.pdf}
    \caption{a density plot of the aggression parameter k at max energy}
    \label{fig:k density}

    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/e_0_density.pdf}
    \caption{a density plot of the aggression parameter $e_0$ at max energy}
    \label{fig:e_0 density}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/k_history.pdf}
    \caption{a line(ish) plot of the aggression parameter $k$ at max energy, the grey bounds are the standard deviation}
    \label{fig:k history}

    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/e_0_history.pdf}
    \caption{a line(ish) plot of the aggression parameter $e_0$  k at max energy the grey bounds are the standard deviation}
    \label{fig:e_0 history}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=15cm,height=8cm,keepaspectratio]{figures/aggression_density.pdf}
    \caption{a density plot of aggression at max energy. I am trying to get it to be multiple plots}
    \label{fig:aggression density}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=15cm,height=10cm,keepaspectratio]{figures/num_winners.pdf}
    \caption{Number of successful males at different patch areas, gaps are because it took longer than 12 hours to run}
    \label{fig:num winners}
\end{figure}

>>>>

Within the immature phase, we assume males grow with rate
\begin{equation} \label{eq:growth}
\frac{{\mathrm d} m}{{\mathrm d} t} = a \, m^ b
\end{equation}
where $a$ is mass-based rate of growth. Integrating eq. \ref{eq:growth}, the size of individuals at time $t$, having started growing at time $0$ is given by
\begin{equation} \label{eq:growth}
m(t) = \left(m(0)^{1-b} + t a(1-b)\right)^{\frac1{1-b}},
\end{equation}

@@

Finally, we assume males transition from immature $\rightarrow$ mature according to logistic probability function
\begin{equation}\label{eq:allocation}
Pr({\mathrm mature} \, | m) = \frac1{1 + \exp\left(\alpha \left(1 - m / m_{\mathrm mat}\right)\right)},
\end{equation}
where $m_{\mathrm mat}$ is the size at which 50\% of the population is mature and $\alpha$ determines the width of the curve.

@@

\clearpage
\subsection{Contest Algorithm}
Contests 
When a searching male encounters a nest that is occupied a contest over possession of that nest occurs. Contests occur instantaneously.

The costs of a contest escalates with more commitment.

We will assume an escalation curve $kappa$ such that:
$$\kappa : Commitment \rightarrow Intensity$$
where $Intensity$ is $\frac{Energy}{Commitment}$

We use the escalation curve $$ \kappa (c) = c^2 $$

At the start of the contest each male chooses how much  energy they are willing to commit.
They do this by reporting their level of ``commitment'' $c_i$.

The male that chooses to commit more of its energy to the contest will win with probability $= logit^(k(c_i - c_j))$. Where k is is a parameter (I reduced it to 10 (also what )).

The energy that this male chose to commit will then be deducted from both males with the loser continuing to search for mates

To get the cost of the fight we take the integral of $\kappa$:

$$Energy Spent = \int_0^{c_i}{\kappa(t)}dt$$

if the $\kappa$ is $commitment^2$ then
$$EnergySpent = \frac{c_i^3}{3}$$
Where $c_i$ is the losing male.

The male chooses how much energy to commit based off of a commitment function, we examine two different commitment functions, for a male $i$ against a male $j$

An exponential one:
$$\mathrm{commitment} = e^\beta \cdot (\frac{m_i}{m_j})^\alpha $$

And a linear one:

$$\mathrm{commitment} = \beta + \alpha \cdot (\frac{m_i - m_j}{m_i + m_j})$$

Where the trait values:
$$\alpha , \beta$$
Are under selection.

@@

    \item aggression 0 ($\beta$) called $e_0$
    \item aggression 1 ($\alpha$) called $k$

@@

\section{Results} % (fold)
\label{sec:results}
% energy expenditure plots
\begin{figure}[h!]
    \centering
    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/energy_pc_plot_exp_f10.pdf}
    \caption{Exponential commitment function, ftime = 10. Percent energy expenditure of cohort at generation 499}
    \label{fig:energy pc}

    \centering
    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/energy_total_plot_exp_f10.pdf}
    \caption{Exponential commitment function, ftime = 10. Total energy expenditure of cohort at generation 499}
    \label{fig:energy total}
\end{figure}
\clearpage

\begin{figure}[h!]
    \centering
    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/energy_pc_plot_exp_f15.pdf}
    \caption{Exponential commitment function, ftime = 15. Percent energy expenditure of cohort at generation 499}
    \label{fig:var energy pc}
    \centering

    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/energy_total_plot_exp_f15.pdf}
    \caption{Exponential commitment function, ftime = 15. Total energy expenditure of cohort at generation 499}
    \label{fig:var energy total}
\end{figure}

% parameter plots density  plots

% exp 10
\begin{figure}[h!]
    \centering
    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f10_k_density.pdf}
    \caption{Exponential commitment function, ftime = 10. a density plot of the aggression parameter k}
    \label{fig:k density}

    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f10_e_0_density.pdf}
    \caption{Exponential commitment function, ftime = 10. a density plot of the aggression parameter $e_0$}
    \label{fig:e_0 density}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f10_k_history.pdf}
    \caption{Exponential commitment function, ftime = 10. $k$ the grey bounds are the standard deviation}
    \label{fig:k history}

    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f10_e_0_history.pdf}
    \caption{ Exponential commitment function, ftime = 10. $e_0$ the grey bounds are the standard deviation}
    \label{fig:e_0 history}
\end{figure}
% exp 15

\begin{figure}[h!]
    \centering
    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f15_k_density.pdf}
    \caption{Exponential commitment function, ftime = 15. a density plot of the aggression parameter k}
    \label{fig:k density}

    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f15_e_0_density.pdf}
    \caption{Exponential commitment function, ftime = 15. a density plot of the aggression parameter $e_0$.}
    \label{fig:e_0 density}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f15_k_history.pdf}
    \caption{Exponential commitment function, ftime = 15. $k$ the grey bounds are the standard deviation}

    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f15_e_0_history.pdf}
    \caption{Exponential commitment function, ftime = 15. $e_0$ the grey bounds are the standard deviation}
\end{figure}

%% WINNERS

\begin{figure}[h!]
    \centering
    \includegraphics[width=15cm,height=10cm,keepaspectratio]{figures/exp_f10_num_winners.pdf}
    \caption{exp ftime = 10, exp commitment Number of successful males at different patch areas}
    \label{fig:num winners expf10}

    \includegraphics[width=15cm,height=10cm,keepaspectratio]{figures/exp_f15_num_winners.pdf}
    \caption{exp ftime = 15, exp commitment. Number of successful males at different patch areas, gaps are extinctions}
    \label{fig:num winners}
\end{figure}

### 2016-02-16 scrap/continuous_escalation.Rmd
fixed bug with cts contests
cts contest theory merged into ms, plots of cts in ms

if the $\kappa$ is $commitment^2$ then:
$$EnergySpent = \frac{c_i^3}{3}$$
Where $c_i$ is the losing male

@@

We define normalised mass difference as so:
$$\bar{M_i} = \frac {M_i} {M_i + M_j}$$

$$\bar{M_j} = \frac {M_j} {M_i + M_j}$$

>>>>

if the $\kappa$ is $commitment^2$ the:n
$$EnergySpent = \frac{c_i^3}{3}$$
Where $c_i$ is the losing male

@@

We define normalised mass difference as so:
$$\bar{M_i} = \frac {M_i} {M_i + M_j}, \bar{M_j} = \frac {M_j} {M_i + M_j}$$
